---
title: "DATA 607 Week1"
author: "Puja Roy"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

### Link to Article: 
https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

### Overview

This data set is about the majors that recent grads from colleges/universities have studied/pursued. It provides information about recent grads including both men and women employment rates, major category, and how many students performed non college or low wage jobs.

### Load R packages

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Retrieve URL of the raw dataset from GitHub

```{r code-chunk-label}
url <- "https://raw.githubusercontent.com/pujaroy280/DATA607Week1/main/recent-grads.csv"
```

### Load the CSV dataset into a data frame

```{r}
data <- read.csv(url) # Data has 173 observations (rows) of 21 variables (columns)

```

### Display the first few rows of the dataset
```{r}
head(data)

```

### Summary statistics
```{r}
summary(data)
```

### Explore the structure of the dataset
```{r}
str(data)
```
```{r}
# Select single column
employed <- data$Employed

# Select multiple columns
jobs_data <- data[, c("College_jobs", "Non_college_jobs", "Low_wage_jobs")]

print(employed)
print(jobs_data)

```

```{r}
# Change values in the Major column to lowercase
data$Major <- tolower(data$Major)
print(data)
```

### Create a scatter plot
```{r}
# Plot with x-axis labels without scientific notation
ggplot(data = data, aes(x = Total, y = Major_category)) + 
  geom_point() +
  scale_x_continuous(labels = function(x) format(x, scientific = FALSE))

```

### Subsetting Data
```{r}
# Data frame is named 'data'
# Subset data to include only rows where the Major_category is engineering
engineering_data <- data[data$Major_category == "Engineering", ]

# Get the count of rows in the subset data frame
num_engineering <- nrow(engineering_data)

# Print the count
print(num_engineering)

```

```{r}
# Data frame is named 'data'
# Subset data to include only rows where the Major_category is education
education_data <- data[data$Major_category == "Education", ]

# Get the count of rows in the subset data frame
num_education <- nrow(education_data)

# Print the count
print(num_education)
```

```{r}
# Data frame is named 'data'
# Subset data to include only rows where the Major_category is Biology & Life Science
biolifesci_data <- data[data$Major_category == "Biology & Life Science", ]

# Get the count of rows in the subset data frame
num_biolifesci <- nrow(biolifesci_data)

# Print the count
print(num_biolifesci)
```

```{r}
# Data frame is named 'data'
# Subset data to include only rows where the Major_category is Arts
arts_data <- data[data$Major_category == "Arts", ]

# Get the count of rows in the subset data frame
num_arts <- nrow(arts_data)

# Print the count
print(num_arts)
```

```{r}
# Data frame is named 'data'
# Subset data to include only rows where the Major_category is Psychology & Social Work
psysocwork_data <- data[data$Major_category == "Psychology & Social Work", ]

# Get the count of rows in the subset data frame
num_psysocwork <- nrow(psysocwork_data)

# Print the count
print(num_psysocwork)
```

```{r}
# Assuming your data frame is named 'data'
# Subset data to include only rows where Major_category is Engineering
engineering_data <- data[data$Major_category == "Engineering", ]

# Sum the Women column to get the total number of women in engineering
num_women_engineering <- sum(engineering_data$Women, na.rm = TRUE)

# Print the count
print(num_women_engineering)

```

```{r}
# Assuming your data frame is named 'data'
# Subset data to include only rows where Major_category is Engineering
engineering_data <- data[data$Major_category == "Engineering", ]

# Sum the Women column to get the total number of women in engineering
num_men_engineering <- sum(engineering_data$Men, na.rm = TRUE)

# Print the count
print(num_men_engineering)

```

### Using ggplot2
```{r}
# Load necessary packages
library(ggplot2)

# Subset data to include only rows where Major_category is Engineering
engineering_data <- data[data$Major_category == "Engineering", ]

# Sum the Women and Men columns separately
sum_women <- sum(engineering_data$Women, na.rm = TRUE)
sum_men <- sum(engineering_data$Men, na.rm = TRUE)

# Create a data frame for plotting
gender_counts <- data.frame(Gender = c("Women", "Men"),
                            Count = c(sum_women, sum_men))

# Plot the counts
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.5, color = "black", size = 3.5) +  # Add text labels
  scale_fill_manual(values = c("lightblue", "pink")) +  # Change fill colors
  labs(title = "Number of Women and Men in Engineering",
       x = "Gender",
       y = "Count") +
  theme_minimal() +
  scale_y_continuous(labels = function(y) format(y, scientific = FALSE))

```

```{r}
# Load necessary packages
library(ggplot2)

# Subset data to include only rows where Major_category is Business
business_data <- data[data$Major_category == "Business", ]

# Sum the Women and Men columns separately
sum_women <- sum(business_data$Women, na.rm = TRUE)
sum_men <- sum(business_data$Men, na.rm = TRUE)

# Create a data frame for plotting
gender_counts <- data.frame(Gender = c("Women", "Men"),
                            Count = c(sum_women, sum_men))

# Plot the counts
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.5, color = "black", size = 3.5) +  # Add text labels
  scale_fill_manual(values = c("lightblue", "pink")) +  # Change fill colors
  labs(title = "Number of Women and Men in Business",
       x = "Gender",
       y = "Count") +
  theme_minimal() +
  scale_y_continuous(labels = function(y) format(y, scientific = FALSE))

```
```{r}
# Load necessary packages
library(ggplot2)

# Subset data to include only rows where Major_category is Education
education_data <- data[data$Major_category == "Education", ]

# Sum the Women and Men columns separately
sum_women <- sum(education_data$Women, na.rm = TRUE)
sum_men <- sum(education_data$Men, na.rm = TRUE)

# Create a data frame for plotting
gender_counts <- data.frame(Gender = c("Women", "Men"),
                            Count = c(sum_women, sum_men))

# Plot the counts
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.5, color = "black", size = 3.5) +  # Add text labels
  scale_fill_manual(values = c("lightblue", "pink")) +  # Change fill colors
  labs(title = "Number of Women and Men in Education",
       x = "Gender",
       y = "Count") +
  theme_minimal() +
  scale_y_continuous(labels = function(y) format(y, scientific = FALSE))

```
```{r}
# Load necessary packages
library(ggplot2)

# Subset data to include only rows where Major_category is Agriculture & Natural Resources
agricnatres_data <- data[data$Major_category == "Agriculture & Natural Resources", ]

# Sum the Women and Men columns separately
sum_women <- sum(agricnatres_data$Women, na.rm = TRUE)
sum_men <- sum(agricnatres_data$Men, na.rm = TRUE)

# Create a data frame for plotting
gender_counts <- data.frame(Gender = c("Women", "Men"),
                            Count = c(sum_women, sum_men))

# Plot the counts
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.5, color = "black", size = 3.5) +  # Add text labels
  scale_fill_manual(values = c("brown", "lightgreen")) +  # Change fill colors
  labs(title = "Number of Women and Men in Agriculture & Natural Resources",
       x = "Gender",
       y = "Count") +
  theme_minimal() +
  scale_y_continuous(labels = function(y) format(y, scientific = FALSE))

```

### Conclusion
Based on the data analysis conducted, I would like to work on analyzing and visualizing the how many men and women pursued majors related to STEM and how many candidates did off campus, low wage or college jobs. 